## Matches
To access matches from these leagues:
```{r}
fem_matches = FreeMatches(fem_leagues)
paged_table(head(fem_matches))
```


```{r}
wsl = fem_leagues %>%
  filter(season_name == "2020/2021")
wsl_matches = FreeMatches(wsl)
dim(wsl_matches)
```

This season in the WSL has 131 matches, due to a match not played between Tottenham and Birmingham City, where the away team (Birmingham) was punished with a 1 point deduction for fielding an ineligible player.

```{r}
colnames(wsl_matches)
```


That said, it is possible to calculate how the league rank finished. But first, some data cleaning is necessary
```{r}
wsl_matches = wsl_matches %>%
  select(c(match_id, match_date, match_week, home_team.home_team_name, away_team.away_team_name, 
           home_score, away_score, stadium.name, referee.name)) %>%
  rename(home_team = home_team.home_team_name, away_team = away_team.away_team_name)

wsl_matches$result = if_else(wsl_matches$home_score > wsl_matches$away_score, 
                                "HOME", 
                                if_else(wsl_matches$away_score > wsl_matches$home_score, "AWAY", "DRAW"))
# Add non disputed match win for Tottenham Hotspur (no goals but +3 points)
wsl_matches[nrow(wsl_matches) + 1,] = list(0, "2021-01-10", 11, "Tottenham Hotspur Women", "Birmingham City WFC", 0, 0, "The Hive Stadium", "Jane Simms", "HOME")
```

### TODO: Matches plots...


### League Ranking
To get the ranking, it is possible to calculate Home/Away rankings and melt both summing the data by the team name.
```{r}
home_rank = wsl_matches %>%
 group_by(home_team) %>%
 summarise(MP = n(), W = sum(result == "HOME"), D = sum(result == "DRAW"), L = sum(result == "AWAY"),
           GF = sum(home_score), GA = sum(away_score), Points = 3*W+D) %>%
 rename(Team = home_team)


away_rank = wsl_matches %>%
 group_by(away_team) %>%
 summarise(MP = n(), W = sum(result == "AWAY"), D = sum(result == "DRAW"), L = sum(result == "HOME"),
           GF = sum(away_score), GA = sum(home_score), Points = 3*W+D) %>%
 rename(Team = away_team)

# Add penalty for Birmingham City (-1 point)
away_rank["Points"][away_rank["Team"]=="Birmingham City WFC"] = away_rank["Points"][away_rank["Team"]=="Birmingham City WFC"] - 1

rank = rbind(home_rank, away_rank)
rank = aggregate(.~Team, rank, sum)
# rank$Points = rank$W * 3 + rank$D
rank$Points_per_match = rank$Points / rank$MP

kable(rank %>%
 arrange(desc(Points_per_match)))
```


